---
title: "PERSONAL PROJECT"
author: "Veronica Pham"
date: "2023-09-12"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# INTRODUCTION:

A pneumonia outbreak with an initially unknown cause was first reported internationally from Wuhan, China, on December 31, 2019. This virus is now commonly known as COVID-19 or, informally, as Coronavirus. Since that time, the global community has been actively combatting this pandemic. Various measures have been implemented to slow the spread of the virus, a practice often referred to as "flattening the curve." Unfortunately, the world has witnessed the loss of many lives due to the pandemic, and social distancing measures have been put in place to mitigate its impact.

## This report will aim to address the following key questions related to the COVID-19 pandemic:

1. **Which countries have had the highest number of deaths due to COVID-19?**
   - This question seeks to identify the countries that have experienced the most fatalities as a result of the virus.

2. **Which countries have had the highest number of positive cases against the number of tests?**
   - This question focuses on countries where the ratio of positive cases to the number of tests conducted is the highest, indicating the severity of the outbreak in those regions.

3. **Which countries have made the best effort in terms of the number of COVID-19 tests conducted related to their population?**
   - Here, we will determine which countries have conducted a substantial number of tests relative to their population size, reflecting their proactive testing efforts.

4. **Which countries were ultimately the most and least affected related to their population?**
   - This question aims to identify the countries that have been significantly impacted by COVID-19 in relation to their population size, helping to assess the extent of the pandemic's effect in different regions.

By analyzing data and statistics, we can provide insights into the countries that have faced the greatest challenges and those that have demonstrated effective strategies in managing the pandemic.


---------

```{r}
# Set a CRAN mirror
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```


```{r setup, include=FALSE}
library(readr)
read.csv("covid19.csv")
```


```{r}
covid_df <- read.csv("covid19.csv")
dim(covid_df)
vector_cols <- colnames(covid_df)
vector_cols
head(covid_df)
```


The "vector_cols" variable will represent a character vector containing the column names of the "covid_df" dataframe. This character vector is essentially a list of strings, where each string represents a column name in the dataframe. The colnames function extracts the column names from the dataframe and stores them as elements in this vector.


```{r}

install.packages("tibble")
library(tibble)

```


```{r}
glimpse(covid_df)
```


Provides a concise summary of the dataset's structure. It displays key information such as the number of observations (rows) and variables (columns) in the dataset.


```{r}
install.packages("dplyr")
library(dplyr)
```


```{r}
covid_df <- covid_df %>% filter(Province_State == "All States")
covid_df <- covid_df %>% select(-Province_State)
covid_df_all_states <- covid_df
```


```{r}
library(dplyr)
selected_columns <- covid_df_all_states %>% select(Date, Continent_Name, Two_Letter_Country_Code, positive, hospitalized, recovered, death, total_tested)
covid_df_all_states_cumulative <- selected_columns
```


```{r}
library(dplyr)
selected_columns2 <- covid_df_all_states %>% select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
covid_df_all_states_daily <- selected_columns2
```


```{r}

library(dplyr)
# Summarize and filter the data
covid_df_all_states_cumulative_max <- covid_df_all_states_cumulative %>%
  group_by(Continent_Name, Two_Letter_Country_Code) %>%
  summarize(Max_Death = max(death, na.rm = TRUE)) %>%
  filter(Max_Death > 0)
print(covid_df_all_states_cumulative_max)

```


```{r}
install.packages("ggplot2")  
library(ggplot2)  # Load the ggplot2 library

# Create the ggplot visualization

# Create a bar chart
ggplot(data = covid_df_all_states_cumulative_max, aes(x = Two_Letter_Country_Code, y = Max_Death, fill = Continent_Name)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  labs(x = "Country Code", y = "Maximum Death Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```


```{r}

# Sort the data by Max_Death in descending order
sorted_data <- covid_df_all_states_cumulative_max %>%
  arrange(desc(Max_Death))

# Select the top three affected countries
top_3_countries <- head(sorted_data$Two_Letter_Country_Code, 3)

# Store the findings in a character vector
death_top_3 <- as.character(top_3_countries)
print(death_top_3)

```


```{r}
# Make sure to have the dplyr package installed and loaded
install.packages("dplyr")
library(dplyr)

# Summarize and arrange the data
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>%
  group_by(Country_Region) %>%
  summarize(
    `tested` = sum(daily_tested, na.rm = TRUE),
    `positive` = sum(daily_positive, na.rm = TRUE),
    `active` = sum(active, na.rm = TRUE),
    `hospitalized` = sum(hospitalizedCurr, na.rm = TRUE)
  ) %>%
  arrange(desc(`tested`))

# display it in data frame
covid_df_all_states_daily_sum
```


```{r}
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
```


```{r}
# Create the countries vector
countries <- covid_top_10$Country_Region

# Create the tested_cases vector
tested_cases <- covid_top_10$tested

# Create the positive_cases vector
positive_cases <- covid_top_10$positive

# Create the active_cases vector
active_cases <- covid_top_10$active

# Create the hospitalized_cases vector
hospitalized_cases <- covid_top_10$hospitalized

# Name the vectors using the names() function
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries

```


```{r}
# Calculate the positive-to-tested ratio
positive_to_tested_ratio <- positive_cases / tested_cases

# Create a named vector
positive_tested <- setNames(positive_to_tested_ratio, countries)

# Sort the named vector in descending order
positive_tested_sorted <- sort(positive_tested, decreasing = TRUE)

# Get the top three countries with the highest ratio
positive_tested_top_3 <- head(positive_tested_sorted, 3)

# Display the top three countries and their ratios
positive_tested_top_3

```


```{r}
# Create the matrix by combining the vectors
covid_mat <- cbind(tested_cases, positive_cases, active_cases, hospitalized_cases)

# Create the population vector
population <- c(331002651, 145934462, 60461826, 1380004385, 84339067, 37742154, 67886011, 25499884, 32971854, 37846611)

# Perform the multiplication and division
covid_mat <- covid_mat * 100 / population

# Display the resulting matrix
covid_mat

```


```{r}
# compute the ranking for each column
tested_cases_rank <- rank(covid_mat[, "tested_cases"])
positive_cases_rank <- rank(covid_mat[, "positive_cases"])
active_cases_rank <- rank(covid_mat[, "active_cases"])
hospitalized_cases_rank <- rank(covid_mat[, "hospitalized_cases"])

# combine the ranking vectors into a matrix 
covid_mat_rank <- rbind(tested_cases_rank, positive_cases_rank, active_cases_rank, hospitalized_cases_rank)
print(covid_mat_rank)

# display the first row
first_row <- covid_mat_rank[1,]
print(first_row)
```


```{r}
# remove first row of matrix
covid_mat_rank_remove_first_row <- covid_mat_rank[-1,]

# compute the sum of the remaining rows using colSum
aggregated_sum <- colSums(covid_mat_rank_remove_first_row)
print(aggregated_sum)
```


```{r}
# Calculate the positive-to-tested ratio
positive_to_tested_ratio <- positive_cases / tested_cases

# Create a named vector
positive_tested <- setNames(positive_to_tested_ratio, countries)

# Sort the named vector in ascending order
positive_tested_sorted <- sort(positive_tested, decreasing = FALSE)

# Get the top three countries with the lowest ratio (best effort)
best_effort_tested_cased_top_3 <- head(positive_tested_sorted, 3)

# Display the top three countries with the best testing effort
best_effort_tested_cased_top_3

```


```{r}
# Calculate the positive-to-tested ratio
positive_to_tested_ratio <- positive_cases / tested_cases

# Create a named vector
positive_tested <- setNames(positive_to_tested_ratio, countries)

# Find the most affected country
most_affected_country <- names(positive_tested[which.max(positive_tested)])

# Find the least affected country
least_affected_country <- names(positive_tested[which.min(positive_tested)])

# Display the most and least affected countries
most_affected_country
least_affected_country
```


# CONCLUSION:

In conclusion, the analysis reveals significant insights into the impact of COVID-19 across various countries:

1. **Highest Number of Deaths**: The United States, Italy, and the United Kingdom (Great Britain and Northern Ireland) are the top three countries with the highest number of COVID-19-related deaths.

2. **Highest Positive Tests**: The United Kingdom leads in the number of positive COVID-19 tests, followed by the United States and Turkey.

3. **Best Testing Efforts**: Australia, India, and Poland stand out for their commendable testing efforts, considering their respective population sizes.

4. **Most and Least Affected**: The United Kingdom has experienced a substantial impact relative to its population, whereas Australia has been relatively less affected in terms of population size.

These findings shed light on the diverse responses and outcomes of countries in managing the COVID-19 pandemic, reflecting the global challenges and successes in combating the virus.